---
title: "Class 1 Survey"
author: "Addisu Tereda"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Import class one survey data from our Github site

    #check the name of variables

```{r}
C1survey <- read.csv("https://github.com/kijohnson/Advanced-Data-Analysis/raw/refs/heads/main/Class%201%20Survey%20Fall%202025%202.csv")
```

2.  number of people who filled out the survey (row:observations) + the
    number of variables in the dataframe (column: variables) \# Row:
    observation #Column: variables [colomns:58, variables:27]

    #Row: observation #Column: variables

    a\.

    ```{r}
    dim(C1survey)[1]

    ```

    n_obs \<- nrow(C1survey)

    b\.

    ```{r}
    dim(C1survey)[2]
    ```

3.   **a. Rename the column variables to something shorter and that is
    descriptive of what the variable is about (for exampl,e *like_dogs*
    for the ‘Do you like dogs?’ question variable) \#**Display updated
    names

    names(C1survey)

    ```{r}
    names(C1survey)<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
    ```

    ##### **b. Display that they are renamed. Hints: You can use either the rename_with function or the names function. The list of names for time is:**

    ```{r}
    colnames(C1survey)
    ```

4.  **Write code to determine and display the number of factor, integer,
    numerical, and character variables there are in the *C1survey*
    dataset. Suggestion: Try to ask ChatGPT for a solution if you are
    unsure. There is a really short answer to this but we want to see
    what students come up with!**

    #Check variable classes

    sapply(C1survey, class)

    #Count how many are each type

    table(sapply(C1survey, class))

    ```{r}

    table(sapply(C1survey, class))

    ```

##### **5. a. Using code, check the *bday* and *bmonth* variables for any unusual or missing values (hint: use table function). If you find any**

##### #Convert to numeric just in case

C1survey$bday <- as.numeric(C1survey$bday)

C1survey$bmonth <- as.numeric(C1survey$bmonth)

```{r}
table(as.numeric(C1survey$bday), useNA = 'ifany')
```

```{r}
table(as.numeric(C1survey$bmonth), useNA = 'ifany')
```

##### **b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and**

Suppose we find days outside 1–31 or months outside 1–12, we set them NA

#####  **c. after you have cleaned up any unusual values, find the median bday and month. \*Note: please make sure that your bday and month variables are numeric before doing statistics**

```{r}
median(C1survey$bday, na.rm = TRUE)
```

```{r}
median(C1survey$bmonth, na.rm = TRUE)
```

##### **6. a. Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30).** 

```{r}
install.packages("pacman")
pacman::p_load(readr, dplyr)
```

```{r}
C1survey <- C1survey %>%
  mutate(bseason = case_when(
    bmonth %in% c(12, 1, 2) ~ "Winter",
    bmonth %in% c(3, 4, 5)  ~ "Spring",
    bmonth %in% c(6, 7, 8)  ~ "Summer",
    bmonth %in% c(9, 10, 11) ~ "Fall",
    TRUE ~ NA_character_))

```

##### **b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.**

```{r}
tab <- table(C1survey$bmonth, C1survey$bseason)
tab
```

##### **c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.**

```{r}
addmargins(tab)
```

##### **7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).**

proportion of students who like both cats and dogs

```{r}
mean(C1survey$like_cats == "Yes" & C1survey$like_dogs == "Yes", na.rm = TRUE)
```
